---
title: "Metrics Evalulation"
author: "KZ"
date: "8/11/2021"
output: html_document
---

```{r setup, include=FALSE}
options(scipen = 1, digits = 2) #set to two decimal 
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, warning = FALSE)
```


```{r cars}
setwd("/Users/xinyi/Documents/GitHub/network_ngc/R")
library('microbenchmark')
source("defn.R")
source("simulate.r")
source("array2mat.r")
source("pldag.set.r")
source("ngc.r")
source("grangerLasso.r")
source("grangerTLasso_v3.r")
source("grangerThrLasso.r")
source("metrics.R")
library(glmnet)
library(gglasso)
library(igraph)
library("xtable")
```

## Data Generation

We first create the skeleton of a random sparse $d \times p \times p$ network, with group and without group. The time to generate the edges without group is listed below:

```{r pressure, echo=TRUE, cache=TRUE, eval=TRUE}
## Creates the skeleton of a random sparse d x p x p network 
## with 0.1 sparsity
set.seed(123)
d <- 2
p <- 16
n <- 10
edge = defn_net(d = d, p = p, n = n, sparsity = 0.1)

set.seed(123)
grp = c(rep(1,8),rep(2,8))
edge_grp = defn_net(d = d, p = p, n = n, grp = grp, sparsity = 0.1)
microbenchmark(defn_net(d = d, p = p, n = n), times=10)
phi_true = array(c((edge[2,,]),(edge[1,,])), dim = c(p,p,d))
phi_true_grp = array(c((edge_grp[2,,]),(edge_grp[1,,])), dim = c(p,p,d))
```

Then, we simulate $n$ iid samples from the Granger causality network with $p = 10$ variables observed over $T = 50$ time points and the standard error equal to $0.2$. $X$ is a $n \times p \times T$ array. The time to generate the data without group is listed below:

```{r, echo=TRUE, eval=TRUE, cache=TRUE}
###############
## Simulate n iid samples from the granger causality network 
## with p variables observed over T time points
## X is a n x p x T array
set.seed(123)
T <- 20
error_sd <- 0.1
X <- simulate_data(n, edge, T = T, error_sd = error_sd)
X_grp <- simulate_data(n, edge_grp, T = T, error_sd = error_sd)
microbenchmark(simulate_data(n, edge, T = T, error_sd = error_sd),
               times=10)
```

## Methods
We will evaluate three methods with and without group effects.

### Regular Lasso
Apply stardard Lasso approach without specifying group with the none group data. The time and metrics are listed below:
```{r, echo=TRUE, eval=TRUE, cache=TRUE}
fit1 = ngc(X, d = d, method = 'regular', typeIerr = 0.02)
microbenchmark(ngc(X, d = d, method = 'regular', typeIerr = 0.02), 
               times = 10)
```
```{r, eval=TRUE, echo = TRUE}
checkEstimation(phi_true, fit1$estMat, method_name = "Lasso")$metrics
```
Then, we also check the results with lamnda selected by cross validation.

```{r, echo=TRUE, eval=TRUE, cache=TRUE}
fit1 = ngc(X, d = d, method = 'regular')
checkEstimation(phi_true, fit1$estMat, method_name = "Lasso")$metrics
```
#### Regular lasso with grouping
Apply standard Lasso approach without specifying group with the none group data. The time and metrics are listed below:
```{r, echo=TRUE, eval=TRUE, cache=TRUE}
group = c(rep(1,8),rep(2,8))
fit12 = ngc(X, d = d, method = 'regular', typeIerr = 0.02, group = group)
microbenchmark(ngc(X, d = d, method = 'regular', typeIerr = 0.02, group = group), 
               times = 10)

checkEstimation(phi_true, fit12$estMat, method_name = "Lasso")$metrics
```
Apply stardard Lasso approach without specifying group with the group data. The time and metrics are listed below:
```{r, echo=TRUE, eval=TRUE, cache=TRUE}
fit12 = ngc(X_grp, d = d, method = 'regular', typeIerr = 0.02, group = group)
microbenchmark(ngc(X_grp, d = d, method = 'regular', typeIerr = 0.02, group = group), 
               times = 10)

checkEstimation(phi_true_grp, fit12$estMat, method_name = "Lasso")$metrics
```

#### Regular lasso with grouping by time
Apply stardard Lasso approach without specifying group with the none group data. The time and metrics are listed below:
```{r, echo=TRUE, eval=TRUE, cache=TRUE}
fit13 = ngc(X, d = d, method = 'regular', typeIerr = 0.02, group = group, groupByTime = TRUE)
microbenchmark(ngc(X, d = d, method = 'regular', typeIerr = 0.02, group = group, groupByTime = TRUE), 
               times = 10)

checkEstimation(phi_true, fit13$estMat, method_name = "Lasso")$metrics

fit132 = ngc(X, d = d, method = 'regular', typeIerr = 0.02, groupByTime = TRUE)
checkEstimation(phi_true, fit132$estMat, method_name = "Lasso")$metrics

```
Apply stardard Lasso approach without specifying group with the group data. The time and metrics are listed below:
```{r, echo=TRUE, eval=TRUE, cache=TRUE}
fit13 = ngc(X_grp, d = d, method = 'regular', typeIerr = 0.02, group = group, groupByTime = TRUE)
microbenchmark(ngc(X_grp, d = d, method = 'regular', typeIerr = 0.02, group = group, groupByTime = TRUE), 
               times = 10)

checkEstimation(phi_true_grp, fit13$estMat, method_name = "Lasso")$metrics

fit132 = ngc(X_grp, d = d, method = 'regular', typeIerr = 0.02, groupByTime = TRUE)
checkEstimation(phi_true, fit132$estMat, method_name = "Lasso")$metrics

```

### Thresholded Lasso
Apply thresholded Lasso approach without specifying group with the none group data. The time and metrics are listed below:
```{r, echo=TRUE, eval=TRUE, cache=TRUE}
fit2 <- ngc(X, d = d, method = 'threshold', typeIerr = 0.02)
microbenchmark(ngc(X, d = d, method = 'threshold', typeIerr = 0.02),
               times = 10)

checkEstimation(phi_true, fit2$estMat, method_name = "Thresholded Lasso")$metrics
```

#### Thresholded Lasso with grouping
Apply thresholded Lasso approach without specifying group with the none group data. The time and metrics are listed below:
```{r, echo=TRUE, eval=TRUE, cache=TRUE}
fit22 = ngc(X, d = d, method = 'threshold', typeIerr = 0.02, group = group)
microbenchmark(ngc(X, d = d, method = 'threshold', typeIerr = 0.02, group = group), 
               times = 10)

checkEstimation(phi_true, fit22$estMat, method_name = "Thresholded Lasso")$metrics
```
Apply thresholded Lasso approach without specifying group with the group data. The time and metrics are listed below:
```{r, echo=TRUE, eval=TRUE, cache=TRUE}
fit22 = ngc(X_grp, d = d, method = 'threshold', typeIerr = 0.02, group = group)
microbenchmark(ngc(X_grp, d = d, method = 'threshold', typeIerr = 0.02, group = group), 
               times = 10)

checkEstimation(phi_true_grp, fit22$estMat, method_name = "Thresholded Lasso")$metrics
```

#### Thresholded Lasso with grouping by time
Apply thresholded Lasso approach without specifying group with the none group data. The time and metrics are listed below:
```{r, echo=TRUE, eval=TRUE, cache=TRUE}
fit23 = ngc(X, d = d, method = 'threshold', typeIerr = 0.02, group = group, groupByTime = TRUE)
microbenchmark(ngc(X, d = d, method = 'threshold', typeIerr = 0.02, group = group, groupByTime = TRUE), 
               times = 10)

checkEstimation(phi_true, fit23$estMat, method_name = "Thresholded Lasso")$metrics

fit232 = ngc(X, d = d, method = 'threshold', typeIerr = 0.02, groupByTime = TRUE)
checkEstimation(phi_true, fit232$estMat, method_name = "Thresholded Lasso")$metrics

```
Apply thresholded Lasso approach without specifying group with the group data. The time and metrics are listed below:
```{r, echo=TRUE, eval=TRUE, cache=TRUE}
fit23 = ngc(X_grp, d = d, method = 'threshold', typeIerr = 0.02, group = group, groupByTime = TRUE)
microbenchmark(ngc(X_grp, d = d, method = 'threshold', typeIerr = 0.02, group = group, groupByTime = TRUE), 
               times = 10)

checkEstimation(phi_true_grp, fit23$estMat, method_name = "Thresholded Lasso")$metrics

fit232 = ngc(X_grp, d = d, method = 'threshold', typeIerr = 0.02, groupByTime = TRUE)
checkEstimation(phi_true, fit232$estMat, method_name = "Thresholded Lasso")$metrics
```

#### Apply thresholded lasso with unkown order
First, we apply thresholded lasso with ungrouped data.
```{r, echo=TRUE, eval=TRUE, cache=TRUE}
fit24 <- ngc(X, method = 'threshold', typeIerr = 0.02)
microbenchmark(ngc(X, method = 'threshold', typeIerr = 0.02),
               times = 10)
fit24$tsOrder
checkEstimation(phi_true, fit24$estMat[,,1:d], method_name = "Thresholded Lasso")$metrics

fit242 <-ngc(X, method = 'threshold', typeIerr = 0.02, group = group)
fit242$tsOrder
microbenchmark(ngc(X, method = 'threshold', typeIerr = 0.02, group = group), 
               times = 10)

checkEstimation(phi_true, fit242$estMat[,,1:d], method_name = "Thresholded Lasso")$metrics

fit243 <-ngc(X, method = 'threshold', typeIerr = 0.02, group = group, groupByTime = TRUE)
fit243$tsOrder
microbenchmark(ngc(X, method = 'threshold', typeIerr = 0.02, group = group, groupByTime = TRUE), 
               times = 10)

checkEstimation(phi_true, fit243$estMat[,,1:d], method_name = "Thresholded Lasso")$metrics

fit243 <-ngc(X, method = 'threshold', typeIerr = 0.02, groupByTime = TRUE)
fit243$tsOrder
checkEstimation(phi_true, fit243$estMat[,,1:d], method_name = "Thresholded Lasso")$metrics
```
Then, we apply thresholded lasso with grouped data.
```{r thres, echo=TRUE, eval=TRUE, cache=TRUE}
fit24 <- ngc(X_grp, method = 'threshold', typeIerr = 0.02)
microbenchmark(ngc(X_grp, method = 'threshold', typeIerr = 0.02),
               times = 10)
fit24$tsOrder
checkEstimation(phi_true_grp, fit24$estMat[,,1:d], method_name = "Thresholded Lasso")$metrics
```


### Truncating Lasso
Apply truncating Lasso approach with/without specifying group with the none group data. The time and metrics are listed below:
First, we apply truncating lasso with ungrouped data.
```{r truncated, echo=TRUE, cache=TRUE}
fit3 <- ngc(X, d=2, method = 'truncate', typeIerr = 0.02)
microbenchmark(ngc(X, d=2, method = 'truncate', typeIerr = 0.02),
               times = 10)
fit3$tsOrder
checkEstimation(phi_true, fit3$estMat[,,1:d], method_name = "Truncate Lasso")$metrics
```
Then, we apply truncating lasso with grouped data.
```{r, echo=TRUE, cache=TRUE}
fit33 <- ngc(X_grp, d=2, method = 'truncate', typeIerr = 0.05)
microbenchmark(ngc(X_grp, d=2, method = 'truncate', typeIerr = 0.05),
               times = 10)
checkEstimation(phi_true_grp, fit33$estMat[,,1:d], method_name = "Truncate Lasso")$metrics

```

Next, we apply truncating Lasso approach without specifying group with the none group data. The time and metrics are listed below:
First, we apply truncating lasso with ungrouped data.
```{r truncated_auto, echo=TRUE, cache=TRUE}
fit34 <- ngc(X, method = 'truncate', typeIerr = 0.02)
microbenchmark(ngc(X, method = 'truncate', typeIerr = 0.02),
               times = 10)
fit34$tsOrder
checkEstimation(phi_true, fit34$estMat[,,dim(fit34$estMat)[3]-seq(d,1,-1)], method_name = "Truncate Lasso")$metrics

```
Then, we apply truncating lasso with grouped data.
```{r, echo=TRUE, cache=TRUE}
fit34 <- ngc(X_grp, method = 'truncate', typeIerr = 0.05)
microbenchmark(ngc(X_grp, method = 'truncate', typeIerr = 0.05),
               times = 10)
fit34$tsOrder
checkEstimation(phi_true_grp, fit34$estMat[,,1:d], method_name = "Truncate Lasso")$metrics
```
